---
output:
  md_document:
    variant: gfm
    preserve_yaml: TRUE
    pandoc_args:
      - "--wrap=preserve"
always_allow_html: true
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../_posts") })
author: "William C. Rowden"
discussion_id: Note-on-Names
excerpt: How researchers have named developmental stages provides a clue to their empirical characteristics.
layout: post
mermaid: false
tags:
  - growing up
title: Note on Names
---

```{r setup, include=FALSE}
library(kableExtra)
library(knitr)
library(tidyverse)

base_dir <- "~/Documents/GitHub/rowdenw.github.io"
base_url <- "/"
fig_path <- "images/"

knitr::opts_knit$set(base.dir = base_dir, base.url = base_url)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.path = fig_path,
                      cache.path = '../_cache/',
                      message=FALSE, warning=FALSE,
                      cache = TRUE) # Change to TRUE for publication.
```

## Compare Stage Names

How researchers have named developmental stages provides a clue to their empirical characteristics. This blog uses Bill Joiner's names, which are intended to emphasize the strengths of each level, listed from earlier to later stages.

As an example of names portraying strengths, those at the Expert stage are strongly motivated to develop subject-matter expertise, because they assume a leader's legitimate power comes from expertise and positional authority. Similarly the Co-Creator stage understands that everything in business and life is interdependent.`[^1]`{=markdown}

 Joiner borrowed the Expert and Achiever names from Bill Torbert. What Joiner calls "Level of Leadership Agility",`[^2]`{=markdown} Torbert calls "Action-Logic" to emphasize the dynamic`[^3]`{=markdown} by which our theory-in-use shapes our actions, perhaps without our awareness.`[^4]`{=markdown} The [Global Leadership Profile](https://www.gla.global/the-glp/levels-of-leadership-development/) re-names two of the original action-logics, from "Individualist" to "Redefining," and from "Strategist" to "Transforming," to better convey their leadership advantages.`[^5]`{=markdown}

```{r Name-Stages, echo=FALSE}
stages_path <- file.path(base_dir, "_data", "Name-Stages.csv")
stages_df <- read.csv(stages_path, stringsAsFactors = FALSE)
stages_df %>%
  mutate(Version = fct_relevel(Version, "Level of Agility", "Action-Logic",
                               "Leadership Maturity", "Ego Development Stage")) %>%
  arrange(Version) %>%
  filter(E > 1) %>%
  pivot_wider(names_from = c(Version, Researcher),
              names_glue = "<pre>{Version}\n({Researcher})</pre>",
              values_from = Name) %>%
  kable(booktabs = TRUE, caption = "Cross-Reference Stage Names",
        escape = FALSE, format = "html") %>%
  pack_rows(index = c("Preconventional" = 2, "Conventional" = 3, "Postconventional" = 3))
```

Susanne Cook-Greuter's stages of Leadership Maturity`[^6]`{=markdown} provide names for each stage that describe their center or view of self. Her earlier Ego Development Stage names were closer to `[Loevinger's labels]({% post_url 2021-03-07-What-Is-Growing-Up %})`{=markdown}.`[^7]`{=markdown}

While stage names provide some descriptions of the process of development, they can also lead to confusion with personality types or management styles.`[^1]`{=markdown} For example, ["Achiever" is the name of one of the CliftonStrengths' themes](https://www.gallup.com/cliftonstrengths/en/252134/achiever-theme.aspx), which is different from the Achiever action-logic or level of agility. Strengths are enduring talents,`[^8]`{=markdown} whereas the nature of development is change over time. Similarly someone with the [Strategic CliftonStrengths theme](https://www.gallup.com/cliftonstrengths/en/252350/strategic-theme.aspx) is not likely to be at the Strategist action-logic. Consequently reading the stage names as adjectives, looking for one that "fits," is not the recommended way to estimate one's own developmental stage.

## Caution About Self-Estimates

> That's the news from Lake Wobegon, where all the women are strong, all the men are good-looking, and all the children are above average.`[^9]`{=markdown}

The reliability of the Global Leadership Profile is checked against the client's self-estimate.`[^5]`{=markdown} Nevetherless my own limited anecdotal experience is that self-estimates are later than the profile for conventional ratings (i.e., Conformer, Expert, and Achiever), and self-estimates are earlier than the profile for postconventional ratings (e.g., Catalyst and Co-Creator). This reminds me of the Dunning–Kruger Effect:`[^10]`{=markdown}

> The miscalibration of the incompetent stems from an error about the self, whereas the miscalibration of the highly competent stems from an error about others.

Consequently people's self-estimates present us with the Arbus and Bosnins Puzzle:`[^k]`{=markdown}

> Those in one caste, the Arbus, always tell the truth; those in the other, the Bosnins, always tell the opposite of the truth.

In other words, the following statements may not provide much information:

* "I Arbu."
* "I'm competent."
* "I'm above-average."
* "I'm late-stage."

## Notes

`[^1]`{=markdown}: Joiner, W.B.; Josephs, S.A. (2006) _Leadership Agility: Five Levels of Mastery for Anticipating and Initiating Change_; 1st edition.; Jossey-Bass: San Francisco; [ISBN 978-0-7879-7913-3](https://isbn.nu/9780787979133).
`[^2]`{=markdown}: Joiner, B. (2014) [Leadership Agility: A Global Imperative](https://changewise.biz/wp-content/uploads/2018/06/Leadership-Agility-Global-Imperative-Joiner.pdf). _[Dialogue](https://dialoguereview.com/leadership-agility-global-imperative/)_ *[DEC 2013/FEB 2014](https://issuu.com/revistabibliodiversidad/docs/dialogue-dec13-issue2?e=7978833/5843640)*.
`[^3]`{=markdown}: Herdman-Barker, E.; Torbert, B. (2011) Generating and Measuring Practical Differences in Leadership Performance at Postconventional Action-Logics. In _The Postconventional Personality: Assessing, Researching, and Theorizing Higher Development_; Pfaffenberger, A.H., Marko, P.W., Combs, A., Eds.; SUNY Press: Albany; [ISBN 978-1-4384-3465-0](https://isbn.nu/9781438434650).
`[^4]`{=markdown}: Torbert, W.R.; Taylor, S.S. (November 18, 2007) Action Inquiry: Interweaving Multiple Qualities of Attention for Timely Action. In _The SAGE handbook of action research: participative inquiry and practice_; Reason, P., Bradbury, H., Eds.; SAGE: London [ISBN 978-1-4462-7114-8](https://isbn.nu/9781446271148).
`[^5]`{=markdown}: Torbert, W.R. (2016) Brief Comparison of Five Developmental Measures: The GLP, the MAP, the LDP, the SOI, and the WUSCT.
`[^6]`{=markdown}: Metcalf, M. (2016) [Council Post: What Is The Path For Leadership Maturity?](https://www.forbes.com/sites/forbescoachescouncil/2016/08/30/what-is-the-path-for-leadership-maturity/)
`[^7]`{=markdown}: Cook-Greuter, S.R. (2013) [Nine Levels Of Increasing Embrace In Ego Development: A Full-Spectrum Theory Of Vertical Growth And Meaning Making](http://www.cook-greuter.com/Cook-Greuter%209%20levels%20paper%20new%201.1'14%2097p%5B1%5D.pdf). *Prepublication version*, 97.
`[^8]`{=markdown}: Buckingham, M.; Clifton, D.O. (2001) _Now, Discover Your Strengths_; 1st edition.; The Free Press: New York, 2001; [ISBN 978-0-7432-0114-8](https://isbn.nu/9780743201148).
`[^9]`{=markdown}: Flatow, I.; Salvatore, J. (2015) [Are ALL Minnesotans Above Average?](https://www.sciencefriday.com/segments/are-all-minnesotans-above-average/) *Science Friday*.
`[^10]`{=markdown}: Kruger, J.; Dunning, D. (2000) Unskilled and Unaware of It: How Difficulties in Recognizing One’s Own Incompetence Lead to Inflated Self-Assessments. *Journal of Personality and Social Psychology* *77*, 1121–34, doi: [10.1037//0022-3514.77.6.1121](https://doi.org/10.1037//0022-3514.77.6.1121).
`[^k]`{=markdown}: Kraitchik, M. (1953) *Mathematical Recreations*; New York, Dover Publications; [ISBN 978-0-486-20163-4](https://isbn.nu/9780486201634).

