---
output:
  md_document:
    variant: gfm
    preserve_yaml: TRUE
    pandoc_args:
      - "--wrap=preserve"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../_drafts") })
author: "William C. Rowden"
date: "2021-03-21"
discussion_id: Profile-Agile-Consultans
excerpt: How can we grow deeper in wisdom and understanding? The experience of an Agile consultant team suggests some ways.
layout: post
mermaid: false
tags:
  - growing up
title: Profile Agile Consultants
---

How can we grow deeper in wisdom and understanding, into ["post logical" thinking]({% post_url 2021-03-14-Grown-Up-Thinking %})? The experience of an Agile consultant team might suggest some ways.

```{r setup, include=FALSE}
library(knitr)
# library(kableExtra)
library(tidyverse)
library("ggthemes")

base_dir <- "~/Documents/GitHub/rowdenw.github.io"
base_url <- "/"
fig_path <- "images/draft-"

knitr::opts_knit$set(base.dir = base_dir, base.url = base_url)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.path = fig_path,
                      cache.path = '../_cache/',
                      message=FALSE, warning=FALSE,
                      cache = FALSE) # Change to TRUE for publication.
```

```{r Early, echo=FALSE}
joiner <- c("Enthusiast", "Operator", "Conformer", "Expert", "Achiever", "Catalyst",
            "Co-Creator", "Synergist")

profile_path <- file.path(base_dir, "_data", "Source-GLP.csv")
profile_df <- read.csv(profile_path, stringsAsFactors = FALSE)

stems_path <- file.path(base_dir, "_data", "Score-GLP.csv")
stems_df <- read.csv(stems_path, stringsAsFactors = FALSE)

team <- profile_df %>%
  select(Profile, ACN, Client, Rating, Modifier) %>%
  arrange(Rating, Modifier) %>%
  filter(ACN == TRUE, Client == TRUE) %>%
  mutate(Member = row_number())

team_stems <- stems_df  %>%
  inner_join(team, by = "Profile") %>%
  arrange(Member, Stage) %>%
  mutate(Description = paste("Team Member", Member)) %>%
  group_by(Member) %>%
  mutate(Ogive = cumsum((Stems)))
stems <- 30
early <- team_stems %>% filter(Member == 1)
ggplot(early, aes(x = Stage, y = Ogive, color = factor(Rating))) + geom_point() +
  geom_smooth() +
  scale_color_manual("Rating",
                     labels = c("6" = "Achiever", "7" = "Catalyst", "8" = "Co-Creator"),
                     values = c("6" = "orange", "7" = "green", "8" = "#008080")) +
  scale_x_continuous(labels = joiner, limits = c(2,9), n.breaks = 9) +
  scale_y_continuous(limits = c(0,stems)) +
  theme_tufte() +
  labs(y = "Cumulative Stems")
```

```{r Late, echo=FALSE}
member_max <- team[which.max(team$Member), "Member"]
late <- team_stems %>% filter(Member == member_max)
ggplot(late, aes(x = Stage, y = Ogive, color = factor(Rating))) + geom_point() +
  geom_smooth() +
  scale_color_manual("Rating",
                     labels = c("6" = "Achiever", "7" = "Catalyst", "8" = "Co-Creator"),
                     values = c("6" = "orange", "7" = "green", "8" = "#008080")) +
  scale_x_continuous(labels = joiner, limits = c(2,9), n.breaks = 9) +
  scale_y_continuous(limits = c(0,stems)) +
  theme_tufte() +
  labs(y = "Cumulative Stems")
```

```{r Early+Late, echo=FALSE}
early_late <- team_stems %>% filter(Member %in% c(1, member_max))
ggplot(early_late, aes(x = Stage, y = Ogive, color=factor(Rating))) + geom_point() +
  geom_smooth() +
  scale_color_manual("Rating",
                     labels = c("6" = "Achiever", "7" = "Catalyst", "8" = "Co-Creator"),
                     values = c("6" = "orange", "7" = "green", "8" = "#008080")) +
  scale_x_continuous(labels = joiner, limits = c(2,9), n.breaks = 9) +
  scale_y_continuous(limits = c(0,stems)) +
  theme_tufte() +
  labs(y = "Cumulative Stems")
```

```{r Score-Stems, echo=FALSE}
ggplot(team_stems, aes(x = Stage, y = Ogive, color = factor(Rating))) + geom_point() +
  geom_smooth() +
  coord_cartesian(xlim=c(2,9), ylim=c(0,stems)) +
  facet_wrap(~ Member) +
  scale_color_manual("Rating",
                     labels = c("6" = "Achiever", "7" = "Catalyst", "8" = "Co-Creator"),
                     values = c("6" = "orange", "7" = "green", "8" = "#008080")) +
  theme_tufte() +
  labs(y = "Cumulative Stems")
```

```{r Deliberately-Develop, echo=FALSE}
private_path <- file.path(base_dir, "_data", "Private.csv")
private_df <- read.csv(private_path, stringsAsFactors = FALSE)
multiple <- private_df %>%
  group_by(Name) %>%
  filter(n() > 1) %>%
  inner_join(profile_df)
initial <- multiple[which.min(multiple$Year), "Profile"]
subsequent <- multiple[which.max(multiple$Year), "Profile"]
comparison <- stems_df  %>%
  filter(Profile %in% c(initial, subsequent)) %>%
  inner_join(profile_df, by = "Profile") %>%
  group_by(Profile) %>%
  mutate(Ogive = cumsum((Stems)))
ggplot(comparison, aes(x = Stage, y = Ogive, color=factor(Rating))) + geom_point() +
  geom_smooth() +
  scale_color_manual("Rating",
                     labels = c("6" = "Achiever", "7" = "Catalyst", "8" = "Co-Creator"),
                     values = c("6" = "orange", "7" = "green", "8" = "#008080")) +
  scale_x_continuous(labels = joiner, limits = c(2,9), n.breaks = 9) +
  scale_y_continuous(limits = c(0,stems)) +
  theme_tufte() +
  labs(y = "Cumulative Stems")
```

## Conclusion

* consultant aspects that encourage growth
* Agile aspects that encourage growth
* horizontal decalage
* feedback as it relates to growth
* deliberate development

## Notes

`[^a]`{=markdown}: Cohn, L.D. (1998) Age trends in personality development: A quantitative review. In _Personality development: Theoretical, empirical, and clinical investigations of Loevinger’s conception of ego development_; Lawrence Erlbaum Associates Publishers: Mahwah, NJ, US; pp. 133–143 [ISBN 978-0-8058-1649-5](https://isbn.nu/9780805816495).
`[^b]`{=markdown}: Cook-Greuter, S.R. (2013) [Nine Levels Of Increasing Embrace In Ego Development: A Full-Spectrum Theory Of Vertical Growth And Meaning Making](http://www.cook-greuter.com/Cook-Greuter%209%20levels%20paper%20new%201.1'14%2097p%5B1%5D.pdf). *Prepublication version*, 97.
`[^c]`{=markdown}: "Co-Creator" is called "Transforming" in Torbert, W.R. (n.d.) [The Seven Levels of Leadership Development and Their Impact](https://www.gla.global/the-glp/levels-of-leadership-development/).
`[^d]`{=markdown}: Kent Beck, et al. (2001) [Manifesto for Agile Software Development](https://agilemanifesto.org/).
`[^e]`{=markdown}: Personal communication with Global Leadership Associates (2017).
`[^f]`{=markdown}: Cook‐Greuter, S.R. (2004) Making the Case for a Developmental Perspective. _Ind and Commercial Training_  *36*, 275–281, doi:[10.1108/00197850410563902](https://doi.org/10.1108/00197850410563902).
`[^g]`{=markdown}: Moen, R.D.; Norman, C.L. (2010) [Circling Back](http://www.apiweb.org/circling-back.pdf). _Quality Progress_ *43*, 22–28.
`[^h]`{=markdown}: Argyris, C.; Schon, D.A. (1974) _Theory in Practice: Increasing Professional Effectiveness_; 1st edition.; Jossey-Bass Publishers: San Francisco; [ISBN 978-0-87589-230-6](https://isbn.nu/9780875892306).
`[^i]`{=markdown}: Including my profile, my estimate matched the profile four times, was mismatched by 0.5 stage five times, was mismatched by 1 stage one time, was mismatched by 1.5 stages two times, and was mismatched by 2 stages one time.

